\section{Platform Architecture}

The observability platform is not a single monitoring tool. It is a distributed system where each
environment produces its own signals: metrics from nodes, workloads, and control-plane components;
logs from pods and hosts; cluster events; and telemetry from storage, networking, Kafka, databases,
and facility services. Those signals are collected close to where they are produced, tagged with
site and cluster context, and made available through shared backends and dashboards.

\subsection{Metrics}

Prometheus handles metric scraping. Instead of maintaining a static list of targets by hand, the
platform uses discovery labels; a service becomes visible to the monitoring system just by following
the expected conventions. The same idea applies to alert rules, probes, and scrape configs: the
platform finds what it should watch by looking for the right labels \cite{prometheus-operator}.

For long-term storage, Mimir serves as Prometheus's scalable backend. This means metrics are not
discarded after a few days; you can compare behavior across nights, clusters, and commissioning
periods. Mimir runs as a distributed service with separate components for ingestion, querying,
compaction, caching, and object-storage access \cite{grafana-mimir}. At this scale, the
observability system is not lightweight. The data it produces needs to be planned for, not just
collected.

\subsection{Logs}

Alloy runs close to the nodes and collects host logs, pod logs, and cluster events. Before sending
anything to Loki, it enriches each log line with metadata: namespace, pod, container, application,
severity \cite{lsst-it-k8s-cookbook,grafana-loki-alloy}. But Kubernetes is not the only source.
Network devices, firewalls, power systems, and other facility equipment connected to the observatory
network also send logs, and those flow into the same pipeline. Loki is also distributed with
gateways, distributors, ingesters, queriers, query frontends, index gateways, compactors, and
caches \cite{lsst-it-k8s-cookbook,grafana-loki}. The end result is one place to search logs across
all clusters and all connected systems, with enough labels to narrow things down quickly.

\subsection{Grafana}

Grafana is where the raw signals turn into actual questions: which cluster is unhealthy, which
namespace is eating resources, whether Ceph is under pressure, whether Kafka latency is climbing,
whether a network device is dropping packets, or whether something in the facility---power,
cooling---is part of the problem. The dashboards are treated as part of the operational model, not
as decoration.

\subsection{Alerting}

Alerts are small contracts: what condition matters, how urgent it is, who should get it, and what
context they need to start debugging. The goal is not to alert on everything. It is to alert when
someone actually needs to do something, and give them enough to move quickly.

\subsection{Discovery}

One of the quieter but more important parts of the platform is how it finds hosts to monitor.
There is no hand-maintained list. Instead, the platform queries the Puppet instances that manage
the observatory's infrastructure. Every time a new host is provisioned through the
infrastructure-as-code process, it becomes visible to the observability platform automatically---no
code changes, no manual additions, no tickets to update a configuration file.

This matters more than it might seem. In an environment that is still growing and changing, a
static host list becomes wrong almost immediately. A node brought up for commissioning, a host
added for a new instrument---all of them appear in the platform on their own, because the source of
truth is the infrastructure itself, not a separate list that someone has to remember to keep
current.

Services follow a different path, using Kubernetes-native discovery through label conventions. But
for the physical and virtual hosts that make up the observatory network, the answer is automatic:
provision the host, and the platform finds it.

\subsection{Controlled Change}

Every configuration change---dashboards, alert rules, retention policies, ingress config, label
selectors, and so on---moves through staged environments before reaching production
\cite{lsst-it-k8s-cookbook}. This means operational knowledge is versioned. A dashboard can be
reviewed, promoted, or rolled back. An alert can be tuned after an incident. A retention policy
can be adjusted when data volume grows. This is especially important during commissioning, when
both the system and the questions asked about it are constantly changing.

The platform does not hide its complexity. It makes it explicit and reproducible, which allows it
to keep improving rather than slowly becoming a pile of manual fixes.

% \textbf{TODO: ADD ARCHITECTURE DIAGRAM HERE}