\section{GitOps}

The platform is managed as code. Every change, a dashboard, an alert rule, the scraping
configuration, a retention policy, goes through a review process before it reaches production.
Nothing is edited directly on the cluster \cite{lsst-it-k8s-cookbook}.

Changes move through staged environments: development, testing, and production. This matters for
an observatory because you need to know not just what changed, but why. The history of every
decision is preserved and reviewable.

\subsection{Dashboards as Code}

Grafana does not natively support managing dashboards through GitOps; dashboards are normally
created and saved through the UI. To solve this, the team built a templating layer that takes
dashboard definitions from the repository and automatically deploys them to the cluster, where
Grafana picks them up \cite{lsst-it-k8s-cookbook}. This means a dashboard can be reviewed before
it goes out, reverted if it breaks something, and understood later by someone who was not there
when it was created. Dashboards are not personal views saved in someone's account; they are
operational artifacts shared by the whole team.

\subsection{Alert Rules as Code}

Prometheus has some built-in support for loading rules from files, but wiring that into a GitOps
workflow with proper staging, review, and promotion is not something that comes out of the box
\cite{lsst-it-k8s-cookbook,prometheus-alerting,prometheus-alertmanager}. The team built the same
kind of templating layer for alert rules, so they follow the same path as everything else. Each
rule carries enough context, a summary, a description, and a receiver, to be useful to whoever
gets paged at night. Rules that are vague or noisy get fixed through the same review process, so
quality improves over time.

\begin{lstlisting}[
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!10},
    rulecolor=\color{gray!50},
    caption={Alert rule for low PVC free space}
]
groups:
  - name: k8s.rules
    rules:
      - alert: PVCLowFreeSpace
        annotations:
          summary: PVC {{ $labels.persistentvolumeclaim }} is low on free space
          description: >
            PVC {{ $labels.persistentvolumeclaim }} in namespace {{ $labels.namespace }}
            has less than 20% free space.
        expr: |
          (kubelet_volume_stats_available_bytes{job="kubelet",metrics_path="/metrics",namespace=~".*"}
          / kubelet_volume_stats_capacity_bytes{job="kubelet",metrics_path="/metrics",namespace=~".*"}) < 0.20
          and kubelet_volume_stats_used_bytes{job="kubelet",metrics_path="/metrics",namespace=~".*"} > 0
        for: 2m
        labels:
          prod: "true"
          severity: warning
          node_name: '{{ $labels.prom_cluster }}'
          device: null
          service_name: null
\end{lstlisting}
